|

### 2020-11-19 manuscript/article.Rmd
add article in RMarkdown

\documentclass[a4paper,12pt]{article}
\usepackage{anysize}
\usepackage[T1]{fontenc}
\usepackage[stable]{footmisc}
\usepackage{setspace}
\usepackage{lmodern}
\usepackage{libertine}
\usepackage[libertine]{newtxmath}
\usepackage[scale=0.825]{FiraMono}
\usepackage[top=2cm,bottom=2cm,left=2cm,right=2cm]{geometry}
\usepackage{mathtools}
\usepackage[authoryear]{natbib}
\usepackage[UKenglish]{babel}
\usepackage[UKenglish]{isodate}
\usepackage{babelbib}
\usepackage{graphicx}
\usepackage{booktabs, makecell, longtable}
\usepackage{dcolumn}
\usepackage{float}
\usepackage[caption = false]{subfig}
\floatplacement{figure}{H}
\usepackage{caption}
\usepackage{rotating}
\usepackage{pdflscape}
\usepackage{pdflscape}
\newcommand{\blandscape}{\begin{landscape}}
\newcommand{\elandscape}{\end{landscape}}
\usepackage{ifthen}
\usepackage{graphicx}
\usepackage{markdown}
\usepackage[usenames,dvipsnames]{xcolor}
\definecolor{darkblue}{rgb}{0.0,0.0,0.55}
\usepackage{tikz}
\usetikzlibrary{shapes.geometric, arrows}
\setcitestyle{aysep={}}
\usepackage{etoolbox}
\makeatletter
\patchcmd{\NAT@citex}
  {\@citea\NAT@hyper@{%
	 \NAT@nmfmt{\NAT@nm}%
	 \hyper@natlinkbreak{\NAT@aysep\NAT@spacechar}{\@citeb\@extra@b@citeb}%
	 \NAT@date}}
  {\@citea\NAT@nmfmt{\NAT@nm}%
   \NAT@aysep\NAT@spacechar\NAT@hyper@{\NAT@date}}{}{}
\patchcmd{\NAT@citex}
  {\@citea\NAT@hyper@{%
	 \NAT@nmfmt{\NAT@nm}%
	 \hyper@natlinkbreak{\NAT@spacechar\NAT@@open\if*#1*\else#1\NAT@spacechar\fi}%
   {\@citeb\@extra@b@citeb}%
	 \NAT@date}}
  {\@citea\NAT@nmfmt{\NAT@nm}%
   \NAT@spacechar\NAT@@open\if*#1*\else#1\NAT@spacechar\fi\NAT@hyper@{\NAT@date}}
  {}{}
\makeatother
\cleanlookdateon
\exhyphenpenalty=1000
\hyphenpenalty=1000
\widowpenalty=1000
\clubpenalty=1000
\usepackage{hyperref}

\hypersetup{
	breaklinks=true,
	linkcolor=Mahogany,
	citecolor=Mahogany,
	urlcolor=darkblue,
	colorlinks=true}

\doublespacing

\title{The Effect of Legislature Size on Public Spending:\\ A Meta-Analysis\thanks{The authors thank Guilherme Duarte, Robert Myles McDonnell, and David Skarbek for their constructive feedback. We also thank Cedric Antunes, Luis Castro, Giovanna França, Julia Oriente, and Lucas Mingardi for their excellent research assistance. Replication materials are available at \url{https://github.com/danilofreire/distributive-politics-meta-analysis}. We kindly acknowledge funding from the São Paulo State Science Foundation (FAPESP grant number 2018/00646-1).}}

\vspace{2cm}

\author{Huzeyfe Alptekin\thanks{Research Associate, Contemporary Brazilian History Research and Documentation Center, School of Social Sciences, Fundação Getulio Vargas, Brazil, \href{mailto:huzeyfealptekin@gmail.com}{\texttt{huzeyfealptekin@gmail.com}}.}
\hspace{.5cm} Danilo Freire\thanks{Independent Researcher, \href{mailto:danilofreire@gmail.com}{\texttt{danilofreire@gmail.com}}, \url{http://danilofreire.github.io}. Corresponding author.} 
\hspace{.5cm} Umberto Mignozzetti\thanks{Visiting Assistant Professor, \href{mailto:umberto.mignozzetti@emory.edu}{\texttt{umberto.mignozzetti@emory.edu}}, \url{http://umbertomig.com}.} 
\hspace{.5cm} Catarina Roman\thanks{Research Associate, School of International Relations, Fundação Getulio Vargas, Brazil, \href{mailto:catarinamroman@gmail.com}{\texttt{catarinamroman@gmail.com}}.}}

\vspace{2cm}
\date{\today}
\begin{document}
\maketitle

\begin{abstract}

\noindent In a seminal article, \citet{weingast1981political} argue that there
is a positive relationship between legislature size and inefficiency in public
expenditures. Their proposition is currently known as the ``law of $1/n$'' and
has been widely debated in political science and public administration.
However, recent studies have questioned the validity of the theory. In this
letter, we estimate the first meta-analysis of the relationship between the
number of legislators and public spending. Based on a sample of \textbf{29}
articles, we find only mixed evidence for the effect of legislature size on
government budgets. Yet the aggregate results mask important differences. While
earlier, non-causal studies provide moderate support for the ``law of $1/n$'',
recent papers using causal inference methods consistently find a negative
relation between seats and spending, suggesting that previous results might
have been driven by spurious correlations or omitted variable bias. The
available evidence also indicates that XXXXXXXXXXXXXXXX

\vspace{.4cm}
\noindent \textbf{Keywords:} distributive politics; law of $1/n$; legislature
size; meta-analysis; public spending

\noindent \textbf{JEL Codes:} H21; H23; H50; H61
\end{abstract}

@@

\section{Introduction}
\label{sec:intro}

Over the past decades, a large literature has examined the relationship between
legislature size and public expenditure. \citet{weingast1981political} provided
the general framework to analyse distributive politics. The authors argue that
the larger the number of legislative districts ($n$), the smaller the share of
tax burden each one will bear ($1/n$), thus legislators have an incentive to
overspend in their districts and transfer the costs to the entire polity.
However, recent studies have questioned the validity of the ``law of $1/n$'', as
the theory is currently known. For instance, spending limits, strong executives,
and bicameralism may reduce the inefficiency of pork barrel projects
\citep{bradbury2009spatially,chen_malhotra_2007,primo2006stop}. Moreover,
\citet{primo2008distributive} affirm that, due to spatial spillovers, a
collection of small districts can supply public goods more efficiently than the
central government. The authors conclude that a ``reverse law of $1/n$'' may
hold, wherein a higher number of legislators in small constituencies decrease
the overall public spending.

@@

In this letter, we conduct the first meta-analysis that tests the generality of
the ``law of $1/n$''. We have selected 29 articles that use quantitative methods
to analyse the impact of legislature size over government spending. We run two
sets of models: the first group uses only the main estimates of each paper (42),
while the second set includes all coefficients reported in our sample
(142).\footnote{Although there are 29 papers in our review, we include 42
coefficients in the restricted sample. This is because 13 articles report
separate estimates for the impact of lower and upper house size on government
expenditures. In the full model, we add every outcome reported in the
papers, regardless of whether the coefficient comes from the main tables or
from robustness checks conducted by the original authors. We discuss our
sampling process in the Methods section of this paper and in the Supplementary
Material.} Our independent variables consist of three measures, namely the size
of the lower chamber ($n$), its natural logarithm (log($n$)), and the number of
members in the upper chamber ($k$). We evaluate their effect upon public
expenditure per capita, the natural logarithm of public expenditure per capita,
and government expenditure as a percentage of GDP. We then estimate if the
results vary according to four potential sources of effect heterogeneity.

Aggregate results show that council size has no significant impact on public
spending. However, the meta-regressions suggest that our study sample has high
levels of heterogeneity, and effects differ substantially according to research
design specifications. Whereas earlier studies employing convention spending
regressions often find a positive relation between legislature size and
government budgets, recent studies that use new methods of causal inference,
specially those with regression discontinuity designs, provide consistent
evidence for a negative relation. This goes in contrast with the original
theory and supports the existence of a ``reverse law of $1/n$''.
\textbf{Moreover, ...}  In summary, our results point out that if legislature
size affects government budgets, they do so under specific conditions and are
mediated by domestic factors.

\section{Methods}

@@

\footnotesize
\begin{longtable}{>{\raggedright\arraybackslash}p{2.9cm}
   >{\raggedright\arraybackslash}p{4.2cm}
   >{\centering\arraybackslash}p{1.3cm}
   >{\centering\arraybackslash}p{1cm}
   >{\centering\arraybackslash}p{1cm}
   >{\centering\arraybackslash}p{1.4cm}
   >{\centering\arraybackslash}p{2.5cm}}
\caption{Papers included in the meta-analysis, listed alphabetically by last name}\\
\toprule
\raggedright Author(s) & \centering Title & \centering Journal & \centering Year & \centering Country & \centering Method & \raggedleft Electoral System
\tabularnewline
\midrule
\endhead
R. Baqir & Districts, Spillovers, and Government Overspending & Unpub. & 1999 & USA & OLS & M\\
R. Baqir & Districting and Government Overspending & JPE & 2002 & USA & OLS & M\\
T. Baskaran & Coalition Governments, Cabinet Size, and the Common Pool Problem: Evidence from the German States & EJPE & 2013 & DEU & IV & NM\\
\begin{tabular}[t]{@{}l@{}}T. Bjedov,\\ S. Lapointe \& \\T. Madiès \end{tabular} & The Impact of Within-Party and Between-Party Ideological Dispersion on Fiscal Outcomes: Evidence from Swiss Cantonal Parliaments & PC & 2014 & CHE & Panel & NM\\
\begin{tabular}[t]{@{}l@{}}J. Bradbury \& \\ M. Crain\end{tabular} & Legislative Organization and Government Spending: Cross-Country Evidence & JPubE & 2001 & Multiple (37) & Panel & NM\\
\begin{tabular}[t]{@{}l@{}}J. Bradbury \& \\ E. F. Stephenson\end{tabular} & Spatially Targeted Government Spending and Heterogeneous Constituent Cost Shares & JPriE & 2009 & USA & OLS & M\\
\begin{tabular}[t]{@{}l@{}}J. Chen \& \\ N. Malhotra\end{tabular} & The Law of 1/K: The Effect of Chamber Size on Government Spending in Bicameral Legislatures & APSR & 2007 & USA & OLS & M\\
\begin{tabular}[t]{@{}l@{}}G. Crowley\end{tabular} & The Law of 1/N Revisited: Distributive Politics, Legislature Size, and the Costs of Collective Action &  SEJ & 2019 & USA & Panel & M\\
\begin{tabular}[t]{@{}l@{}}J. Drew \& \\ B. Dollery\end{tabular} & The Price of Democracy? Political Representation Structure and Per Capita Expenditure in Victorian Local Government & UAR & 2017 & AUS & Panel & NM\\
\begin{tabular}[t]{@{}l@{}}H. A. Erler\end{tabular} & Legislative Term Limits and State Spending & PC & 2007 & USA & Panel & M\\
\begin{tabular}[t]{@{}l@{}}N. Fiorino \& \\ R. Ricciuti\end{tabular} & Legislature Size and Government Spending in Italian Regions: Forecasting the Effects of a Reform & PC & 2007 & ITA & IV & NM\\
\begin{tabular}[t]{@{}l@{}}D. Höhmann\end{tabular} & The Effect of Legislature Size on Public Spending: Evidence from a Regression Discontinuity Design & PC & 2017 & DEU & RDD & NM\\
\begin{tabular}[t]{@{}l@{}}A. Kessler\end{tabular} & Communication in Federal Politics: Universalism, Policy Uniformity, and the Optimal Allocation of Fiscal Authority & JPE & 2014 & USA & Panel & M\\
\begin{tabular}[t]{@{}l@{}}D. Lee\end{tabular} & Supermajority Rule and Bicameral Bargaining & PC & 2016 & USA & Panel & M\\
\begin{tabular}[t]{@{}l@{}}D. Lee \& \\ S. Park\end{tabular} & Court-Ordered Redistricting and the Law of 1/N & PC & 2018 & USA & Panel & M\\
\begin{tabular}[t]{@{}l@{}}B. Lewis\end{tabular} & Legislature Size, Local Government Expenditure and Taxation, and Public Service Access in Indonesia & SCID & 2019 & IDN & RDD & NM\\
\begin{tabular}[t]{@{}l@{}}V. Lledo\end{tabular} & Electoral Systems, Legislative Fragmentation and Public Spending: A Comparative Analysis of Brazilian States & Unpub. & 2003 & BRA & OLS & NM\\
\begin{tabular}[t]{@{}l@{}}L. MacDonald\end{tabular} & The Impact of Government Structure on Local Public Expenditures & PC & 2008 & USA & OLS & M\\
\begin{tabular}[t]{@{}l@{}}B. Maldonado\end{tabular} & Legislature, Leaders, and Leviathans: How Constitutional Institutions Affect the Size of Government Spending & SSQ & 2013 & Multiple (92) & OLS & NM\\
\begin{tabular}[t]{@{}l@{}}J. Matsusaka\end{tabular} & The Endogeneity of the Initiative: A Comment on Marschall and Ruhil & SPPQ & 2005 & USA & IV & M\\
\begin{tabular}[t]{@{}l@{}}B. Mukherjee\end{tabular} & Political Parties and the Size of Government in Multiparty Legislatures: Examining Cross-Country and Panel Data Evidence & CPS & 2003 & Multiple (110) & Panel & NM\\
\begin{tabular}[t]{@{}l@{}}P. Pettersson-Lidbom\end{tabular} & Does the Size of the Legislature Affect the Size of Government? Evidence from Two Natural Experiments & JPubE & 2012 & FIN \& SWE & RDD & NM\\
\begin{tabular}[t]{@{}l@{}}D. Primo\end{tabular} & Stop Us before We Spend Again: Institutional Constraints on Government Spending & E\&P & 2006 & USA & Panel & M\\
\begin{tabular}[t]{@{}l@{}}R. Ricciuti\end{tabular} & Legislature Size, Bicameralism and Government Spending: Evidence from Democratic Countries & Unpub. & 2004 & Multiple (75) & OLS & NM\\
\begin{tabular}[t]{@{}l@{}}C. Schaltegger \& \\ L. Feld\end{tabular} & Do Large Cabinets Favor Large Governments? Evidence on the Fiscal Commons Problems for Swiss Cantons & JPubE & 2009 & CHE & Panel & NM\\
\begin{tabular}[t]{@{}l@{}}E. Stein,\\ E. Talvi \& \\ A. Grisanti\end{tabular} & Institutional Arrangements and Fiscal Performance: The Latin American Experience & Unpub. & 1998 & Multiple (26) & OLS & NM\\ [1ex]
\bottomrule
\label{tab:papers}
\begin{minipage}{\textwidth}
\renewcommand{\footnoterule}{}
\vspace{-0.5cm}
\footnotetext{\textbf{Note:} Unpub$=$Unpublished, JPE$=$Journal of Political
Economy, EJPE$=$European Journal of Political Economy, PC$=$Public Choice,
JPubE$=$Journal of Public Economics, JPriE$=$Journal of Private Enterprise,
APSR$=$American Political Science Review, SEJ$=$Southern Economic Journal,
UAR$=$Urban Affairs Review, SCID$=$Studies in Comparative International
Development, SSQ$=$Social Science Quarterly, SPPQ$=$State Politics and Policy
Quarterly, CPS$=$Comparative Political Studies, E\&P$=$Economics and Politics,
OLS$=$Ordinary least squares, IV$=$Instrumental variables, Panel$=$Panel
data/fixed effects, RDD$=$Regression discontinuity design, M$=$Majoritarian,
NM$=$Non-majoritarian (mixed or proportional representation). Country codes
follow the ISO 3166-1 alpha-3 international standard.}
\end{minipage}
\end{longtable}

\normalsize

Our study sample reflects the development of the literature. Although the ``law
of $1/n$'' was first formulated in 1981, the empirical assessment of the theory
only started a few years later, as dates of publishing range from 1998 to 2019.

@@

We also coded four moderators that may help us understand the heterogeneity in
the reported results. We included them in our meta-regressions alongside an
indicator for the type of independent variable used in the original study ($n$,
log($n$), or $k$). The additional moderators are: 1) publication year; 2) paper
publication in an academic journal; 3) electoral system; 4) estimation method.
Since the literature on the ``law of $1/n$'' is notably diverse, we were
restricted to include only the moderators that either refer to important
theoretical questions, such as the effect of the electoral system on public
spending, or to essential characteristics of the publications themselves.
Although more moderators exist in the literature (ie., data aggregation level),
they do not appear as often as required for the meta-regressions. Table
\ref{tab:descriptive} shows the descriptive statistics of the moderator
variables.

@@

\begin{table}[htpb]
 \caption{Descriptive Statistics of Moderators}
 \centering
 \begin{tabular}{lccc}
\hline  \hline
& Main Coefficients & Other Coefficients &   All Coefficients    \\ 
&        N=36        &       N=92        &    N=128    \\ 
\hline     
Independent Variables: &             &                    &                  \\
$\qquad$ $k$ &    9 (25.0\%)      &     29 (31.5\%)    & 38 (29.7\%)  \\ 
$\qquad$ $n$ &      21 (58.3\%)     &    51 (55.4\%)   & 72 (56.2\%)  \\ 
$\qquad$ log($n$) &      6 (16.7\%)     &     12 (13.0\%)    & 18 (14.1\%)  \\ 
Year &     2009 (6.50)    &    2008 (5.75)   & 2008 (5.98) \\ 
Published work: &             &                    &                  \\ 
$\qquad$Yes &     30 (83.3\%)     &    74 (80.4\%)    & 104 (81.2\%) \\ 
$\qquad$No &    6 (16.7\%)       &     18 (19.6\%)   & 24 (18.8\%)  \\ 
Electoral system: &             &                    &                  \\ 
$\qquad$Majoritarian &     18 (50.0\%)      &    40 (43.5\%)   & 58 (45.3\%)  \\ 
$\qquad$Non-Majoritarian &    18 (50.0\%)       &    52 (56.5\%)   & 70 (54.7\%)  \\ 
Estimation method: &             &                    &                  \\ 
$\qquad$OLS &     13 (36.1\%)     &   43 (46.7\%)     & 56 (43.8\%)  \\ 
$\qquad$Panel &     17 (47.2\%)     &    40 (43.5\%)    & 57 (44.5\%)  \\ 
$\qquad$IV &     3 (8.33\%)      &     3 (3.26\%)    &  6 (4.69\%)  \\ 
$\qquad$RDD &     3 (8.33\%)      &     6 (6.52\%)    &  9 (7.03\%)  \\ 
\hline  
\end{tabular}
\label{tab:descriptive}
\begin{minipage}{\textwidth}
\renewcommand{\footnoterule}{}
\footnotetext{\textbf{Note:} The values for the variable \textit{Year} indicate the mean and the standard deviation, respectively.}
\end{minipage}    
\end{table}

@@

\section{Results}

\subsection{Binomial Z-Tests}
\label{sub:binomial}

The ``law of $1/n$'' states that more legislators increase government expenditure.
Here we employ three methods to assess the empirical validity of that
relationship. First, we run a binomial one-proportion z-test to determine
whether the study coefficients have a positive or negative sign. Then, we fit 7
random-effect models using the \texttt{meta} \citep{balduzzi2019perform} and the
\texttt{dmetar} \citep{dmetar2019} packages for the \texttt{R} statistical
language \citep{rstats2019}. We estimate the true effect size variance
($\tau^2$) with a Restricted Maximum Likelihood Estimator, which the literature
considers the most precise for continuous dependent variables
\citep{veroniki2016methods}. Lastly, we run a series of meta-regressions to test
possible sources of result heterogeneity. To recapitulate, our independent
variables of interest are lower chamber size ($n$), the natural logarithm of
lower chamber size (log($n$)), and upper chamber size ($k$). The dependent
variables are public expenditure per capita, the natural logarithm of public
expenditure per capita, and government expenditure as a percentage of GDP. Since
the outcomes have different scales, we treat them separately in our models.

The binomial z-test evaluates whether the coefficient of our independent
variables are positive or negative. The null hypothesis here states that the
sign of the coefficient is equally likely to be positive or negative. We start
with $n$, lower house size. Our results indicate that there is no correlation
between the number of legislators in the lower house and public expenditure 
(successes = 10, trials = 21, $p_{success} = 0.476$, 95\% CI = [0.257; 0.702],
$p$-value = 1). Note that the ``law of $1/n$'' suggests that there is a positive
association between both. The binomial test for log($n$) also shows a
non-statistically significant result (successes = 5, trials = 6, $p_{success} =
0.833$, 95\% CI = [0.359; 0.996], $p$-value = 0.219). In contrast, we find a
positive result for the number of legislators in the upper house ($k$), which is
in line with the mainstream literature (successes = 8, trials = 9, $p_{success}
= 0.889$, 95\% CI = [0.518; 0.997], $p$-value = 0.039).

\subsection{Meta-Analysis}
\label{sub:Meta-Analysis}

@@

In our restricted sample, larger lower house size also does not increase
government spending as a percentage of GDP (studies = 5, SMD = -0.008, SE =
0.013, 95\% CI = [-0.045; 0.0285], $p$-value = 0.567, prediction interval =
[-0.105; 0.089], $I^2$ = 87.1\%). We find a positive effect in our extended
sample and the coefficient is statistically significant at the 10\% level
(coefficients = 21, SMD = 0.008, SE = 0.004, 95\% CI = [-0.000; 0.016], $p$-value =
0.058, prediction interval = [-0.026; 0.042], $I^2$ = 98.5\%). The 
result supports the ``law of 1/$n$'', but the evidence is moderate
as the prediction interval include zero.

@@

In a nutshell, we find only weak evidence in favour of the ``law of $1/n$''.
While some models do show a positive and statistically significant result, none
of the prediction intervals are totally positive or negative. The studies also
have considerable heterogeneity, what indicates that the original coefficients
do not point consistently towards the same direction.

\subsection{Meta-Regressions}
\label{sub:regressions}

@@

\section{Discussion}
\label{sec:discussion}

In this article, we assess the empirical validity of the ``law of $1/n$''. Based
on a sample of 26 recent publications on the topic, our meta-analyses show that
there is little evidence that an increase in the number of legislators has a
significant effect on public expenditures. If such effect exists, it is likely
driven by an increase in $k$, the size of the upper legislature, as suggested by
several studies in the literature \citep[e.g.,][]{baqir2002districting, 
bradbury2001legislative, bradbury2003local, chen_malhotra_2007,
gilligan2001fiscal, primo2006stop}. We find no robust evidence for the ``reverse
law of $1/n$'', which posits that larger legislatures lead to lower government
spending. One possible explanation is that the two logics are balancing each
other, thus leading us to a null impact of legislature size on public
expenditure.

@@

Our analyses suggest three areas for further research. First, our study sample
did not include articles that evaluate the association between the log($n$) and
public expenditure per capita or between $k$ and log expenditure per capita.
New work on that area might clarify some of the inconsistencies we find
here. Second, despite the inclusion of several moderators in our models,
aggregate results still show considerable heterogeneity. That is, much of the
disparities among studies are yet to be explained. Domestic factors such as
party dynamics or gerrymandering \citep{lee2015supermajority,
mukherjee2003politicalparties, gilligan2006public} may prove useful in this
regard. Finally, we highlight the need for more evidence in favour or against
the ``law of $1/n$'' to inform general public policies. The available empirical
evidence points out that contextual variables may either amplify or reduce the
impact of larger legislatures on government budgets. They should be taken into
account if policy-makers want to reach an optimal balance between sound fiscal
policy and the demands for increased political representation.

\newpage
\bibliography{references.bib}
\bibliographystyle{apalike}

@@

\end{document}

>>>>

---
title:  |
        | The Effect of Legislature Size on Public Spending
        | A Meta-Analysis
author: 
  - Huzeyfe Alptekin[^huzeyfe]
  - Danilo Freire[^danilo]
  - Umberto Mignozzetti[^umberto]
  - Catarina Roman[^catarina]
date: \today
thanks: "The authors thank Guilherme Duarte, Robert Myles McDonnell, and David Skarbek for their constructive feedback. We also thank Cedric Antunes, Luis Castro, Giovanna França, Julia Oriente, and Lucas Mingardi for their excellent research assistance. Replication materials are available at <https://github.com/danilofreire/distributive-politics-meta-analysis>. We kindly acknowledge funding from the São Paulo State Science Foundation (FAPESP grant number 2018/00646-1)."
abstract: "In a seminal article, @weingast1981political argue that there is a positive relationship between legislature size and inefficiency in public expenditures. Their proposition is currently known as the law of $1/n$ and has been widely debated in political science and public administration. However, recent studies have questioned the validity of the theory. In this letter, we estimate the first meta-analysis of the relationship between the number of legislators and public spending. Based on a sample of 29 articles, we find only mixed evidence for the effect of legislature size on government budgets. Yet the aggregate results mask important differences. While earlier, non-causal studies provide moderate support for the law of $1/n$, recent papers using causal inference methods consistently find a negative relation between seats and spending, suggesting that previous results might have been driven by spurious correlations or omitted variable bias. The available evidence also indicates that XXXXXXXXXXXXXXXX"
abstractspacing: onehalf
keywords: distributive politics; law of $1/n$; legislature size; meta-analysis; public spending
jelcodes: H21; H23; H50; H61
fontsize: 12pt
margin: 2cm
urlcolor: darkblue
linkcolor: Mahogany
citecolor: Mahogany
spacing: double
papersize: a4paper
bibliography: references.bib
biblio-style: apalike
output:
  pdf_document:
    citation_package: natbib
    fig_caption: yes
    number_sections: yes
    keep_tex: no
    toc: no
    toc_depth: 3
    template: article-template.latex
---

```{r, message=FALSE, warning=FALSE, echo=FALSE}
# If you need to install any package while knitting the document
r <- getOption("repos")
r["CRAN"] <- "https://cran.rstudio.com/"
options(repos = r)
if (!require("kableExtra")) {
    install.packages("kableExtra")
}
```

[^huzeyfe]: Research Associate, Contemporary Brazilian History Research and Documentation Center, School of Social Sciences, Fundação Getulio Vargas, <huzeyfealptekin@gmail.com>.

[^danilo]: Independent Researcher, <danilofreire@gmail.com>, <https://danilofreire.github.io>. Corresponding author.

[^umberto]: Visiting Assistant Professor, Quantitative Theory and Methods Department, Emory University, <umberto.mignozzetti@emory.edu>, <http://umbertomig.com>.

[^catarina]: Research Associate, School of International Relations, Fundação Getulio Vargas, <catarinamroman@gmail.com>, <http://catarinaroman.github.io>.

@@

# Introduction
\label{sec:intro}

Over the past decades, a large literature has examined the relationship between
legislature size and public expenditure. \citet{weingast1981political} provided
the general framework to analyse distributive politics. The authors argue that
the larger the number of legislative districts ($n$), the smaller the share of
tax burden each one will bear ($1/n$), thus legislators have an incentive to
overspend in their districts and transfer the costs to the entire polity.
However, recent studies have questioned the validity of the "law of $1/n$", as
the theory is currently known. For instance, spending limits, strong executives,
and bicameralism may reduce the inefficiency of pork barrel projects
\citep{bradbury2009spatially,chen_malhotra_2007,primo2006stop}. Moreover,
\citet{primo2008distributive} affirm that, due to spatial spillovers, a
collection of small districts can supply public goods more efficiently than the
central government. The authors conclude that a "reverse law of $1/n$" may
hold, wherein a higher number of legislators in small constituencies decrease
the overall public spending.

@@

In this letter, we conduct the first meta-analysis that tests the generality of
the "law of $1/n$". We have selected 29 articles that use quantitative methods
to analyse the impact of legislature size over government spending. We run two
sets of models: the first group uses only the main estimates of each paper (42),
while the second set includes all coefficients reported in our sample
(142).\footnote{Although there are 29 papers in our review, we include 42
coefficients in the restricted sample. This is because 13 articles report
separate estimates for the impact of lower and upper house size on government
expenditures. In the full model, we add every outcome reported in the
papers, regardless of whether the coefficient comes from the main tables or
from robustness checks conducted by the original authors. We discuss our
sampling process in the Methods section of this paper and in the Supplementary
Material.} Our independent variables consist of three measures, namely the size
of the lower chamber ($n$), its natural logarithm (log($n$)), and the number of
members in the upper chamber ($k$). We evaluate their effect upon public
expenditure per capita, the natural logarithm of public expenditure per capita,
and government expenditure as a percentage of GDP. We then estimate if the
results vary according to four potential sources of effect heterogeneity.

Aggregate results show that council size has no significant impact on public
spending. However, the meta-regressions suggest that our study sample has high
levels of heterogeneity, and effects differ substantially according to research
design specifications. Whereas earlier studies employing convention spending
regressions often find a positive relation between legislature size and
government budgets, recent studies that use new methods of causal inference,
specially those with regression discontinuity designs, provide consistent
evidence for a negative relation. This goes in contrast with the original
theory and supports the existence of a "reverse law of $1/n$".
\textbf{Moreover, ...}  In summary, our results point out that if legislature
size affects government budgets, they do so under specific conditions and are
mediated by domestic factors.

# Methods

@@

Our study sample reflects the development of the literature. Although the "law
of $1/n$" was first formulated in 1981, the empirical assessment of the theory
only started a few years later, as dates of publishing range from 1998 to 2019.

@@

We also coded four moderators that may help us understand the heterogeneity in
the reported results. We included them in our meta-regressions alongside an
indicator for the type of independent variable used in the original study ($n$,
log($n$), or $k$). The additional moderators are: 1) publication year; 2) paper
publication in an academic journal; 3) electoral system; 4) estimation method.
Since the literature on the "law of $1/n$" is notably diverse, we were
restricted to include only the moderators that either refer to important
theoretical questions, such as the effect of the electoral system on public
spending, or to essential characteristics of the publications themselves.
Although more moderators exist in the literature (ie., data aggregation level),
they do not appear as often as required for the meta-regressions. Table
\ref{tab:descriptive} shows the descriptive statistics of the moderator
variables.

@@

@@

# Results

## Binomial Z-Tests
\label{sub:binomial}

The "law of $1/n$" states that more legislators increase government expenditure.
Here we employ three methods to assess the empirical validity of that
relationship. First, we run a binomial one-proportion z-test to determine
whether the study coefficients have a positive or negative sign. Then, we fit 7
random-effect models using the \texttt{meta} \citep{balduzzi2019perform} and the
\texttt{dmetar} \citep{dmetar2019} packages for the \texttt{R} statistical
language \citep{rstats2019}. We estimate the true effect size variance
($\tau^2$) with a Restricted Maximum Likelihood Estimator, which the literature
considers the most precise for continuous dependent variables
\citep{veroniki2016methods}. Lastly, we run a series of meta-regressions to test
possible sources of result heterogeneity. To recapitulate, our independent
variables of interest are lower chamber size ($n$), the natural logarithm of
lower chamber size (log($n$)), and upper chamber size ($k$). The dependent
variables are public expenditure per capita, the natural logarithm of public
expenditure per capita, and government expenditure as a percentage of GDP. Since
the outcomes have different scales, we treat them separately in our models.

The binomial z-test evaluates whether the coefficient of our independent
variables are positive or negative. The null hypothesis here states that the
sign of the coefficient is equally likely to be positive or negative. We start
with $n$, lower house size. Our results indicate that there is no correlation
between the number of legislators in the lower house and public expenditure 
(successes = 10, trials = 21, $p_{success} = 0.476$, 95\% CI = [0.257; 0.702],
$p$-value = 1). Note that the "law of $1/n$" suggests that there is a positive
association between both. The binomial test for log($n$) also shows a
non-statistically significant result (successes = 5, trials = 6, $p_{success} =
0.833$, 95\% CI = [0.359; 0.996], $p$-value = 0.219). In contrast, we find a
positive result for the number of legislators in the upper house ($k$), which is
in line with the mainstream literature (successes = 8, trials = 9, $p_{success}
= 0.889$, 95\% CI = [0.518; 0.997], $p$-value = 0.039).

## Meta-Analysis
\label{sub:Meta-Analysis}

@@

In our restricted sample, larger lower house size also does not increase
government spending as a percentage of GDP (studies = 5, SMD = -0.008, SE =
0.013, 95\% CI = [-0.045; 0.0285], $p$-value = 0.567, prediction interval =
[-0.105; 0.089], $I^2$ = 87.1\%). We find a positive effect in our extended
sample and the coefficient is statistically significant at the 10\% level
(coefficients = 21, SMD = 0.008, SE = 0.004, 95\% CI = [-0.000; 0.016], $p$-value =
0.058, prediction interval = [-0.026; 0.042], $I^2$ = 98.5\%). The 
result supports the "law of 1/$n$", but the evidence is moderate
as the prediction interval include zero.

@@

In a nutshell, we find only weak evidence in favour of the "law of $1/n$".
While some models do show a positive and statistically significant result, none
of the prediction intervals are totally positive or negative. The studies also
have considerable heterogeneity, what indicates that the original coefficients
do not point consistently towards the same direction.

## Meta-Regressions
\label{sub:regressions}

@@

# Discussion
\label{sec:discussion}

In this article, we assess the empirical validity of the "law of $1/n$". Based
on a sample of 26 recent publications on the topic, our meta-analyses show that
there is little evidence that an increase in the number of legislators has a
significant effect on public expenditures. If such effect exists, it is likely
driven by an increase in $k$, the size of the upper legislature, as suggested
by several studies in the literature \citep[e.g.,][]{baqir2002districting,
bradbury2001legislative, bradbury2003local, chen_malhotra_2007,
gilligan2001fiscal, primo2006stop}. We find no robust evidence for the "reverse
law of $1/n$", which posits that larger legislatures lead to lower government
spending. One possible explanation is that the two logics are balancing each
other, thus leading us to a null impact of legislature size on public
expenditure.

@@

Our analyses suggest three areas for further research. First, our study sample
did not include articles that evaluate the association between the log($n$) and
public expenditure per capita or between $k$ and log expenditure per capita.
New work on that area might clarify some of the inconsistencies we find
here. Second, despite the inclusion of several moderators in our models,
aggregate results still show considerable heterogeneity. That is, much of the
disparities among studies are yet to be explained. Domestic factors such as
party dynamics or gerrymandering \citep{lee2015supermajority,
mukherjee2003politicalparties, gilligan2006public} may prove useful in this
regard. Finally, we highlight the need for more evidence in favour or against
the "law of $1/n$" to inform general public policies. The available empirical
evidence points out that contextual variables may either amplify or reduce the
impact of larger legislatures on government budgets. They should be taken into
account if policy-makers want to reach an optimal balance between sound fiscal
policy and the demands for increased political representation.

\newpage
\setlength{\parindent}{0cm}
\setlength{\parskip}{5pt}

@@
